---
title: "Advanced Usage of the Model-Free Knockoff Filter"
author: "Matteo Sesia (msesia@stanford.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage of the Model-Free Knockoff Filter}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

The function `MFKnockoffs.filter` is a wrapper around several simpler functions that

1. Construct knockoff variables (various functions with prefix `MFKnockoffs.create`)
2. Compute the test statistic $W$ (various functions with prefix `MFKnockoffs.stat`)
3. Compute the threshold for variable selection (`MFKnockoffs.threshold`)

These functions may be called directly if desired. The purpose of this vignette is to illustrate the flexibility of this package with some examples.

```{r, results='hide', message=FALSE, warning=FALSE}
set.seed(1234)
library(MFKnockoffs)
```

Creating an artificial problem
------------------------------
Let us begin by creating some synthetic data.

```{r}
# Problem parameters
n = 1000         # number of observations
p = 1000         # number of variables
k = 60           # number of variables with nonzero coefficients
amplitude = 7.5  # signal amplitude (for noise level = 1)

# Generate the variables from a multivariate normal distribution
mu = rep(0,p); Sigma = diag(p)
X = matrix(rnorm(n*p),n)

# Generate the response from a logistic model and encode it as a factor.
nonzero = sample(p, k)
beta = amplitude * (1:p %in% nonzero) / sqrt(n)
invlogit <- function(x) exp(x) / (1+exp(x))
y.sample <- function(x) rbinom(n, prob=invlogit(x %*% beta), size=1)
y = factor(y.sample(X), levels=c(0,1), labels=c("A","B"))
```

Looking inside the knockoff filter
----------------------------------
Instead of using `MFKnockoffs.filter` directly, we can run the filter manually
by calling its main components one by one.

The first step is to generate the knockoff variables for the true Gaussian distribution of the variables.
```{r}
X_k = MFKnockoffs.create.gaussian(X, mu, Sigma)
```

Then, we compute the knockoff statistics using 10-fold cross-validated lasso
```{r, results='hide', message=FALSE, warning=FALSE}
W = MFKnockoffs.stat.glmnet_coef_difference(X, X_k, y, nfolds=10, family="binomial")
```

Now we can compute the rejection threshold
```{r}
thres = MFKnockoffs.threshold(W, q=0.15, method='knockoff+')
```

The final step is to select the variables
```{r}
selected = which(W >= thres)
print(selected)
```

The false discovery proportion is
```{r}
fdp <- function(selected) sum(beta[selected] == 0) / max(1, length(selected))
fdp(selected)
```

Performing numerical simulations
--------------------------------
We show how to manually run the knockoff filter multiple times and compute
average quantities. This is particularly useful to estimate the FDR
(or the power) for a particular configuration of the knockoff filter
on artificial problems.
```{r}
# Optimize the parameters needed for generating Gaussian knockoffs, 
# by solving as SDP to minimize correlations with the original variables.
# This calculation requires only the model parameters mu and Sigma, 
# not the observed variables X. Therefore, there is no reason to perform it
# more than once for our simulation.

diag_s = MFKnockoffs.knocks.solve_sdp(Sigma)

# Compute the fdp over 20 iterations
nIterations = 20
fdp_list = sapply(1:nIterations, function(it) {
    # Run the knockoff filter manually, using the pre-computed value of diag_s
    X_k = MFKnockoffs.create.gaussian(X, mu, Sigma, diag_s=diag_s)
    W = MFKnockoffs.stat.glmnet_lambda_signed_max(X, X_k, y, family="binomial")
    t = MFKnockoffs.threshold(W, q=0.15, method='knockoff+')
    selected = which(W >= t)
    # Compute and store the fdp
    fdp(selected)
  })
# Estimate the FDR
mean(fdp_list)
```

See also
--------
If you want to see some basic usage of the knockoff filter, see the [introductory vignette](MFKnockoffs.html).
If you want to see how to use [the original knockoff filter](https://web.stanford.edu/~candes/Knockoffs/), see the [fixed-design vignette](fixed.html).