---
title: "Using the Knockoff Filter with a Fixed Design Matrix"
author: "Matteo Sesia (msesia@stanford.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the Knockoff Filter with a Fixed Design Matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
  
  The `MFKnockoffs` package can also be used to perform controlled variable selection with a fixed design matrix, assuming a linear regression model for the response. In this sense, `MFKnockoffs` is a superset of [the original knockoffs package](https://web.stanford.edu/~candes/Knockoffs/). 

```{r, results='hide', message=FALSE, warning=FALSE}
set.seed(1234)
```

```{r}
# Problem parameters
n = 1000          # number of observations
p = 300           # number of variables
k = 30            # number of variables with nonzero coefficients
amplitude = 4.5   # signal amplitude (for noise level = 1)

# Generate the variables from a multivariate normal distribution
mu = rep(0,p); Sigma = diag(p)
X = matrix(rnorm(n*p),n)

# Generate the response from a linear model
nonzero = sample(p, k)
beta = amplitude * (1:p %in% nonzero) / sqrt(n)
y.sample <- function(X) X %*% beta + rnorm(n)
y = y.sample(X)
```

First examples
--------------
In order to create fixed-design knockoffs, we call `MFKnockoffs.filter` with the parameter `statistic` equal to `MFKnockoffs.stat.glmnet_lambda_difference`. Moreover, since not all statistics are valid with fixed-design knockoffs, we use `MFKnockoffs.stat.glmnet_lambda_difference` instead of the default one (which is based on cross-validation).
```{r, results='hide', message=FALSE}
library(MFKnockoffs)
result = MFKnockoffs.filter(X, y, knockoffs = MFKnockoffs.create.fixed, statistic = MFKnockoffs.stat.glmnet_lambda_difference)
```
We can display the results with
```{r}
print(result)
```
The default value for the target false discovery rate is 0.1. In this experiment the false discovery proportion is
```{r}
fdp <- function(selected) sum(beta[selected] == 0) / max(1, length(selected))
fdp(result$selected)
```

See also
--------
If you want to see some basic usage of the knockoff filter, see the [introductory vignette](MFKnockoffs.html).
If you want to look inside the knockoff filter, see the [advanced vignette](advanced.html).